---
title: "Portfolio 1- Tidy Tuesday Project"
---

> description of the goal: The main goal of this Tidy Tuesday project is to explore and analyze automotive fuel efficiency using the “mpg” dataset. The project aims to:

- Understand the relationship between engine displacement and average fuel efficiency.

- Investigate patterns in the distribution of fuel efficiency across different classes of vehicles.

- Evaluate manufacturer-specific fuel efficiency trends and categorize vehicles into high and low efficiency categories based on average mileage.


```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)
glimpse(mpg)

```



```{r data cleaning}
# Data cleaning and preparation
mpg_clean <- mpg %>%
  filter(!is.na(hwy), !is.na(cty)) %>% # Remove missing values
  mutate(avg_mpg = (hwy + cty)/2)      # Create a new column for average mileage

```


```{r explore relationships}
# Explore relationships: Does engine displacement affect mileage?
mpg_summary <- mpg_clean %>%
  group_by(class, manufacturer) %>%
  summarize(mean_mpg = mean(avg_mpg),
            mean_displ = mean(displ),
            .groups = "drop") %>%
  arrange(desc(mean_mpg))

# based on the mpg_summary, the data is sorted by descending average mileage (mean_mpg), reveals lower average engine displacement (mean_displ) among vehicles with higher mileage.

```


```{r data visualization}
# Data visualization: Mileage vs Engine displacement, by Vehicle Class
ggplot(mpg_clean, aes(x = displ, y = avg_mpg, color = class)) +
  geom_point(alpha = 0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.8) +
  facet_wrap(~class, scales = "free") +
  labs(
    title = "Relationship between Engine Displacement and Mileage by Vehicle Class",
    x = "Engine Displacement (L)",
    y = "Average Mileage (mpg)",
    color = "Vehicle Class",
    caption = "Data Source: ggplot2 mpg dataset"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.caption = element_text(size = 8, color = "gray")
  )
  
```



