---
title: "Portfolio 1- Tidy Tuesday Project"
---

> description of the goal: The main goal of this Tidy Tuesday project is to explore and analyze automotive fuel efficiency using the “mpg” dataset. Through various data transformations, visualizations, and summaries, the project aims to:

- Understand the relationship between engine displacement and average fuel efficiency in different classes of vehicles.

- Investigate patterns in the distribution of fuel efficiency across different classes of vehicles.

- Evaluate manufacturer-specific fuel efficiency trends and categorize vehicles into high and low efficiency categories based on average mileage.


```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)
glimpse(mpg)

```



```{r data cleaning}
# Data cleaning and preparation
mpg_clean <- mpg %>%
  filter(!is.na(hwy), !is.na(cty)) %>% # Remove potential missing values
  mutate(avg_mpg = (hwy + cty)/2)      # Create a new column for average mileage


```


```{r explore relationships}
# Explore relationships: Does engine displacement (displ) affect mileage?
mpg_summary <- mpg_clean %>%
  group_by(class, manufacturer) %>%
  summarize(mean_mpg = mean(avg_mpg),
            mean_displ = mean(displ),
            .groups = "drop") %>%
  arrange(desc(mean_mpg))
```


```{r data visualization}
# Data visualization: Mileage vs Engine displacement, by Vehicle Class
ggplot(mpg_clean, aes(x = displ, y = avg_mpg, color = class)) +
  geom_point(alpha = 0.7, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.8) +
  facet_wrap(~class, scales = "free") +
  labs(
    title = "Relationship between Engine Displacement and Mileage by Vehicle Class",
    x = "Engine Displacement (L)",
    y = "Average Mileage (mpg)",
    color = "Vehicle Class",
    caption = "Data Source: ggplot2 mpg dataset"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.caption = element_text(size = 8, color = "gray40")
  )

  
```



