---
title: "Portfolio 7- Survival Analysis using lung cancer patients"
date: "2025-05-01"
output: html_document
---
> This project explore how patient characteristics— age, sex, and performance status (ph.ecog)—influence survival time in lung cancer patients. ph.ecog is a strong, statistically significant predictor:Patients with worse physical condition had a 63%–70% higher risk of death per unit increase. Sex and age were not significant predictors in the interaction model, though sex alone was significant.

```{r}
# install.packages("survminer")
library(survival)
library(survminer) 

head(lung)
str(lung)

#Create a survival object
lung$surv_obj <- with(lung, Surv(time, status == 2))
# by sex
fit_sex <- survfit(surv_obj ~ sex, data = lung)

# install.packages("xfun")
ggsurvplot(
  fit_sex,
  data = lung,
  pval = TRUE,             
  conf.int = TRUE,          
  risk.table = TRUE,  
  legend.labs = c("Male", "Female"),
  xlab = "Days",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Survival Curve by Sex",
  palette = "Dark2"
)
# Females consistently show higher survival probabilities than males (green line) across the entire follow-up period.
# Visually, the median survival time is: Females: Around ~450 days; Males: Around ~300 days
# 50% of female patients survived longer than ~450 days, whereas for males, 50% died before ~300 days.

```

## Cox proportional hazards model
```{r}
cox_model <- coxph(surv_obj ~ age + sex + ph.ecog, data = lung)
summary(cox_model)
# Being female is associated with a 42.5% lower hazard of death compared to males.
# Age does not reach significance at p < .05

cox_interaction <- coxph(surv_obj ~ age * sex + ph.ecog, data = lung)
summary(cox_interaction)
# Each 1-point increase in ECOG score (worse) increases hazard of death by 63.6%.
# the effect of age on survival does not meaningfully differ by sex.

```

