---
title: "Portfolio 8- How vitamin C affect tooth growth in guinea pigs."
date: "2025-05-01"
output: html_document
---

> Well... I'm out of ideas, but this topic sounds interesting. How does Vitamin C affect the growth of guinea pig teeth? lol. 
> This project explore how Vitamin C dose levels and supplement type (orange juice vs. ascorbic acid) affect tooth growth in guinea pigs. The goal is to understand whether the form of vitamin C delivery influences outcomes, and whether this effect depends on the dose administered.


```{r}
library(ggplot2)
data("ToothGrowth") 

head(ToothGrowth)

# Convert dose to factor for ANOVA
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Tooth length by supplement type and dose
ggplot(ToothGrowth, aes(x = dose, y = len, fill = supp)) +
  geom_boxplot() +
  labs(
    title = "Tooth Growth by Dose and Supplement Type",
    x = "Vitamin C Dose (mg/day)",
    y = "Tooth Length",
    fill = "Supplement"
  ) +
  theme_minimal()

# Tooth growth increases with higher doses for both OJ and VC.
# OJ results in greater tooth growth at low and medium doses (0.5, 1.0 mg/day).
# At the highest dose (2.0 mg/day), VC catches up or slightly surpasses OJ.


```
## T-test
```{r}
subset_0.5 <- subset(ToothGrowth, dose == 0.5)
subset_1   <- subset(ToothGrowth, dose == 1)
subset_2   <- subset(ToothGrowth, dose == 2)

t.test(len ~ supp, data = subset_0.5) # There is a statistically significant difference in tooth growth between OJ and VC at the 0.5 mg dose, wirh OJ leads to significantly more tooth growth than VC 
t.test(len ~ supp, data = subset_1) # OJ again leads to significantly greater tooth length;   
t.test(len ~ supp, data = subset_2) # no significant difference   

```
## ANOVA
```{r}
anova_model <- aov(len ~ supp * dose, data = ToothGrowth)
summary(anova_model)
# the type of supplement affects tooth growth.
# dose level strongly affects growth.
# supplement's effect varies by dose.

TukeyHSD(anova_model)
# On average, OJ leads to significantly more tooth growth than VC across all doses.

# Orange juice leads to more tooth growth than pure vitamin C at lower doses.
# At 2.0 mg/day, both delivery methods are equally effective.

# Pretty much what I expected. After all, there is vc in orange juice as well, along with other trace minerals. I wonder if they would see a difference if they isolated the vc in the orange juice. I think it would have.
# Probably should have used vc and water or other placebo.

```


