---
title: "Portfolio 10- Analyze UFO Sightings"
---

> I got the data from NUFORC (National UFO Reporting Center；https://nuforc.org/databank/). So credit to every dedicated person who has reported the suspected UFO sighting. In the dataset of UFO, I have the reported ufo sightings locations, and the reported ufo shapes, and their respective numbers.
But honestly, most of the UFO sightings they describe are a bit of a pipe dream. The descriptions sound like drones or aircraft.

## Locations
```{r}
library(readxl)
library(dplyr)
library(ggplot2)

ufo <- read_excel("~/Downloads/UFO.xlsx")
colnames(ufo) <- c("location", "location_count", "shape", "shape_count")

# Top 10 locations by report count
top_locations <- ufo %>%
  distinct(location, location_count) %>%
  arrange(desc(location_count)) %>%
  slice_head(n = 10)
# No.1 California (16785); No.2 Florida (8643); No.3 Washington (7448)

ggplot(top_locations, aes(x = reorder(location, location_count), y = location_count)) +
  geom_col(fill = "lightblue") +
  coord_flip() +
  labs(title = "Top 10 Locations for UFO Sightings",
       x = "Location", y = "Number of Reports")

```

## Shape
```{r}
# Top 10 shapes by shapes count
top_shapes <- ufo %>%
  distinct(shape, shape_count) %>%
  arrange(desc(shape_count)) %>%
  slice_head(n = 10)

ggplot(top_shapes, aes(x = reorder(shape, shape_count), y = shape_count)) +
  geom_col(fill = "darkorchid") +
  coord_flip() +
  labs(title = "Top 10 Reported UFO Shapes",
       x = "Shape", y = "Number of Reports")

```

## plot both variables
```{r}
library(tidyr)

# Combine both
long_data <- ufo %>%
  select(location, location_count, shape, shape_count) %>%
  pivot_longer(cols = c(location_count, shape_count),
               names_to = "type", values_to = "count")

long_data <- long_data %>%
  filter((type == "location_count" & location %in% top_locations$location) |
         (type == "shape_count" & shape %in% top_shapes$shape))

ggplot(long_data, aes(x = count, y = reorder(ifelse(type == "location_count", location, shape), count), fill = type)) +
  geom_col() +
  facet_wrap(~ type, scales = "free_y") +
  labs(title = "UFO Sightings by Location and Shape",
       x = "Report Count", y = NULL) 

```

